% theorems_summary.tex
% ---------------------------------------------------------------
% A 2-page summary of the extended-ZHY results, aimed at
% Professor Nakahiro Yoshida. Mirrors the framing of
% thesis_overview_2026_v12 (2016 thesis results + new 2026
% directions + Lean 4 formalisation).
%
% Compile: pdflatex theorems_summary.tex
% ---------------------------------------------------------------
\documentclass[11pt]{article}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{microtype}
\usepackage[hidelinks]{hyperref}

\setlength{\parskip}{4pt plus 1pt}
\setlength{\parindent}{0pt}

\newtheorem{theorem}{Theorem}

\title{\vspace{-1.5em}Exact finite-sample variance and BLUE weights\\
for the weighted Hayashi--Yoshida covariance estimator}
\author{Dr Yang Azzollini \\
\small Oxford DPhil Statistics, 2016 \\
\small \texttt{yang.maths@gmail.com}}
\date{April 2026}

\begin{document}
\maketitle
\vspace{-2em}

\section*{Setup}

Let $X, Y$ be a bivariate Brownian motion on $[0, 1]$ with constant
parameters $\sigma_1^2$, $\sigma_2^2$, $\sigma_{12}$. Observations
$\{X_{t_i^X}\}_{i=0}^m$, $\{Y_{t_j^Y}\}_{j=0}^n$ arrive at fixed
(but arbitrary) timestamps, independently of the price process.
Let $R_i = X_{t_i^X} - X_{t_{i-1}^X}$, $S_j = Y_{t_j^Y} - Y_{t_{j-1}^Y}$
denote the returns, and $\tau_{ij} = |J_i^X \cap J_j^Y|$,
$\tau_i^X = |J_i^X|$, $\tau_j^Y = |J_j^Y|$ the overlap structure.
Write $\Theta = \sum_{ij} \tau_{ij}^2 / (\tau_i^X \tau_j^Y)$.

The \emph{extended weighted Hayashi--Yoshida class} of estimators of
$\sigma_{12}$ is $\big\{\,\hat\sigma_{12}(c) = \sum_{ij} c_{ij} R_i S_j
: c_{ij} \in \mathbb{R}\,\big\}$, with $c_{ij} = 0$ whenever
$\tau_{ij} = 0$. Unbiasedness requires $\sum_{ij} c_{ij} \tau_{ij} = 1$.
The \emph{normalised weighted ZHY estimator} is
$\hat\sigma_{12}(w) = \sum_{ij} w_{ij} R_i S_j / \sum_{ij} w_{ij} \tau_{ij}$.

\section*{Results from the 2016 DPhil thesis (Chapter 5)}

\begin{theorem}[Exact finite-sample variance, thesis Theorem 5.1]
For any weight matrix $w$,
\[
\operatorname{Var}\!\left(\hat\sigma_{12}(w) \,\middle|\, \tau\right)
\;=\; \sigma_1^2 \sigma_2^2 \, A_1(w) \;+\; \sigma_{12}^2 \, A_2(w),
\]
where $A_1(w)$ and $A_2(w)$ are explicit functions of
$\{\tau_{ij}, \tau_i^X, \tau_j^Y\}$. The identity is exact, holds
for any fixed timestamp structure, and requires no distributional
assumption on the observation times beyond their independence from
prices.
\end{theorem}

The proof of Theorem~5.1 on p.~108 of the thesis uses the following
geometric fact, sometimes referred to separately as the
\emph{Interval-Structure Lemma}: for $i \ne k$ and $j \ne l$, the
overlaps $\tau_{il}$ and $\tau_{kj}$ cannot both be strictly
positive, because $J_i^X, J_k^X$ are disjoint consecutive
$X$-intervals and similarly for $Y$.

\textbf{A$_1$-optimal weights (thesis \S 5.4).}
Minimising $A_1(w)$ subject to the unbiasedness constraint gives
$w^*_{ij} = \tau_{ij} / (\tau_i^X \tau_j^Y)$, and at this choice
$A_1(w^*) = 1/\Theta$. Theorem~5.2 of the thesis extends these
expressions to the microstructure-noise model.

\section*{New results (2026)}

\begin{theorem}[Full-variance optimum, general $\kappa$]
Let $\kappa = \sigma_{12}^2 / (\sigma_1^2 \sigma_2^2) \in [0, 1]$.
Minimising $\sigma_1^2 \sigma_2^2 A_1(w) + \sigma_{12}^2 A_2(w)$
under $\sum_{ij} w_{ij} \tau_{ij} = 1$ is a convex quadratic
program whose KKT conditions reduce to an $(m + n + 1) \times
(m + n + 1)$ linear system in the multipliers. The solution
$w^*(\kappa)$ collapses to the thesis A$_1$-optimal weights at
$\kappa = 0$. Numerical study gives a variance reduction of less
than 1\% over the A$_1$-optimal weights for all $\rho \in [0,1]$.
\end{theorem}

\begin{theorem}[Two-step plug-in]
The estimator
$\hat\sigma_{12}^{(2)} = \sum_{ij} w_{ij}^*(\hat\kappa) R_i S_j
/ \sum_{ij} w_{ij}^*(\hat\kappa) \tau_{ij}$, with
$\hat\kappa$ constructed from a consistent first-step $\hat\sigma_{12}^{(1)}$,
is unbiased and attains the fixed-$\kappa$ minimum of
Theorem~2 in the large-sample limit.
\end{theorem}

\begin{theorem}[Cram\'er--Rao match at $\rho = 0$]
At $\rho = 0$, the A$_1$-optimal estimator attains the Cram\'er--Rao
lower bound:
\[
\operatorname{Var}\!\big(\hat\sigma_{12}(w^*) \,\big|\, \tau\big)\Big|_{\rho=0}
\;=\; \frac{\sigma_1^2 \sigma_2^2}{\Theta}
\;=\; I(\sigma_{12} \mid \tau)^{-1}\big|_{\sigma_{12} = 0},
\]
where $I(\sigma_{12} \mid \tau)$ is the Fisher information for
$\sigma_{12}$ conditional on the timestamp structure.
\end{theorem}

\section*{Lean 4 formalisation (2026)}

The 2016 thesis results on the extended-ZHY class and the new
results above have been formalised in the Lean 4 proof assistant
with Mathlib. The development consists of four modules and
contains zero \texttt{sorry} statements. The probabilistic input
is packaged as a single structure recording the Brownian
covariance identities and the Gaussian (Isserlis) fourth-moment
formula; everything downstream is algebraic.
\begin{itemize}\itemsep2pt
\item \texttt{ZHY\_Core}: unbiasedness characterisation, A$_1$-optimality,
Cram\'er--Rao equality at $\rho = 0$.
\item \texttt{ZHY\_BLUE}: $A_1(c) \ge 1 / \Theta$ for every unbiased $c$
(Cauchy--Schwarz); equality at $w^*$.
\item \texttt{ZHY\_Algebra}: combinatorial scaffolding (four
\texttt{Finset.sum} identities) used by the variance decomposition.
\item \texttt{ZHY\_Variance}: exact finite-sample variance
decomposition (Theorem~1 above), including the
Interval-Structure Lemma as a hypothesis on $\tau$.
\end{itemize}

\section*{Positioning with respect to the literature}

To the author's best knowledge, no published work contains (i)~an
exact, finite-sample, timestamp-conditional variance formula for
the weighted Hayashi--Yoshida class, (ii)~a BLUE result for the
class in finite samples, or (iii)~a machine-checked proof of
either. The existing literature ---
Hayashi--Yoshida (2005, 2008, 2011),
Griffin--Oomen (2011),
Bibinger (2014),
Koike (2014--2026),
Ogihara--Yoshida (2012) --- works in the $\lambda \to \infty$
asymptotic regime or with renewal / kernel approximations.

Before submitting, the author would welcome a brief sanity check
from an authority on the Hayashi--Yoshida estimator that these
finite-sample exact-variance and BLUE results have not appeared
elsewhere. Full proofs of Theorem~1 (from the 2016 thesis) and
Theorems~2--4 are available on request, as is the Lean~4 source.

\vspace{0.5em}\hrule\vspace{0.3em}
\footnotesize
Repository (public): \texttt{github.com/ChubbyFaceEuler/Paper\_2026}.
DPhil thesis: \emph{Correlation methods in the statistical analysis of
financial trading data}, University of Oxford, 2016.

\end{document}
